An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-17 EconGrowth.Rnw
Now including a biblography file allowing citations. Used Stern and Kander as a first example.
Now includes a test for the nlmrt package from https://stat.ethz.ch/pipermail/r-help/2012-November/330554.html.

@@

<<CES functions, eval=TRUE>>=
####################
# Returns an nls ces model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadData(countryAbbrev)
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  # Establish default guesses for sigma, gamma_E, beta, lambda_L and lambda_E.
  sigmaGuess <- 0.663 # 0.663 for Sweden
  gamma_EGuess <- 0.01 # Minimum contribution by energy is the default.
  betaGuess <- 0.7 # 0.7 for Sweden
  lambda_LGuess <- 0.011 # 0.011 for Sweden
  lambda_EGuess <- 0.018 # 0.018 for Sweden
  lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-Inf, lambda_E=-Inf)
  upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= Inf, lambda_E= Inf)
  if (countryAbbrev == "US"){
    gamma_EGuess <- 0.147 # 0.147 for Sweden
  }
  if (countryAbbrev == "UK"){
    sigmaGuess <- 0.5
    gamma_EGuess <- 0.147 # 0.147 for Sweden
    betaGuess <- 0.9
    lambda_LGuess <- 0.0
    lambda_EGuess <- 0.0
    lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-0.3, lambda_E=-0.3)
    upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= 0.3, lambda_E= 0.3)
  }
  if (countryAbbrev == "CN"){

  }
  # Runs a non-linear least squares fit to the data with constraints
#   modelCES <- nls(iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
#                                                   * iLabor^beta
#                                                   * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
#                           + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) 
#                                                   * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0)), 
#                   data=dataTable,
#                   start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
#                              lambda_L=lambda_LGuess, lambda_E=lambda_EGuess),
#                   control=nls.control(maxiter=10000, tol=1e-05, minFactor=1/1024, 
#                                       printEval=TRUE, warnOnly=TRUE),
#                   algorithm = "port",
#                   lower = lower,
#                   upper = upper
#                   )
  model <- iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
                                                  * iLabor^beta
                                                  * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
                          + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) 
                                                  * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0))
  modelCES <- nlxb(model,
                   data=dataTable,
                   start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
                              lambda_L=lambda_LGuess, lambda_E=lambda_EGuess)
                   )
print(modelCES)
  return(modelCES)
}
print("US Q")
print(summary(cesModel("US", "Q")))

@@

The CES model has the form

@@

\bibliographystyle{model2-names}
\bibliography{<your-bib-database>}

>>>>

<<nlmrt tests, eval=TRUE>>=
# From https://stat.ethz.ch/pipermail/r-help/2012-November/330554.html
Laptop_sale <- c(1405,1863,2027,2669,2938,5275,6595,6943,8621,10905,12420,22400,32380,31600,34900,43163,47838,47592)
CuSale <- Laptop_sale
time <- seq_along(Laptop_sale)
Bass.Model <- Laptop_sale ~ m * ((p + q)^2/p) * (exp(-(p + q) * time)/((q /p) * exp(-(p + q) * time) + 1)^2)
Bass.Fit <- nls(formula = Bass.Model, start = c(p = 0.03, q = 0.4, m = max(CuSale)), trace = FALSE)
Bass.Fit
library(nlmrt)
xdata<-data.frame(Laptop_sale, time)
Bass.Fitn <- nlxb(formula = Bass.Model, start = c(p = 0.03, q = 0.4, m = max(CuSale)), data=xdata,trace=FALSE)
print("")
print("After nlxb fit:")
Bass.Fitn
print(summary(Bass.Fitn))

Bass.Fitn <- wrapnls(formula = Bass.Model, start = c(p = 0.03, q = 0.4, m = max(CuSale)), data=xdata,trace=FALSE)
print("")
print("After wrapnls fit:")
print(summary(Bass.Fitn))

@@

<<ces functions, eval=FALSE>>=
####################
# Returns an nls ces model for the country and energyType specified.
# energyType should be one of Q", "X", or "U".
##
cesModel <- function(countryAbbrev, energyType){
  # Load the data that we need.
  dataTable <- loadData(countryAbbrev)
  # We need to do the CES fit with the desired energyType.
  # To achieve the correct fit, we'll change the name of the desired column
  # to "iEToFit" and use "iEToFit" in the nls function.
  dataTable <- replaceColName(dataTable, energyType, "iEToFit")
  # Establish default guesses for sigma, gamma_E, beta, lambda_L and lambda_E.
  sigmaGuess <- 0.663 # 0.663 for Sweden
  gamma_EGuess <- 0.01 # Minimum contribution by energy is the default.
  betaGuess <- 0.7 # 0.7 for Sweden
  lambda_LGuess <- 0.011 # 0.011 for Sweden
  lambda_EGuess <- 0.018 # 0.018 for Sweden
  lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-Inf, lambda_E=-Inf)
  upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= Inf, lambda_E= Inf)
  if (countryAbbrev == "US"){
    gamma_EGuess <- 0.147 # 0.147 for Sweden
  }
  if (countryAbbrev == "UK"){
    sigmaGuess <- 0.5
    gamma_EGuess <- 0.147 # 0.147 for Sweden
    betaGuess <- 0.9
    lambda_LGuess <- 0.0
    lambda_EGuess <- 0.0
    lower = list(sigma=-Inf, gamma_E=0.0, beta=-Inf, lambda_L=-0.3, lambda_E=-0.3)
    upper = list(sigma= Inf, gamma_E=1.0, beta= Inf, lambda_L= 0.3, lambda_E= 0.3)
  }
  if (countryAbbrev == "CN"){

  }
  # Runs a non-linear least squares fit to the data with constraints
#   modelCES <- nls(iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
#                                                   * iLabor^beta
#                                                   * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
#                           + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) 
#                                                   * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0)), 
#                   data=dataTable,
#                   start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
#                              lambda_L=lambda_LGuess, lambda_E=lambda_EGuess),
#                   control=nls.control(maxiter=10000, tol=1e-05, minFactor=1/1024, 
#                                       printEval=TRUE, warnOnly=TRUE),
#                   algorithm = "port",
#                   lower = lower,
#                   upper = upper
#                   )
  model <- iGDP ~ ((1-gamma_E)^(1.0/sigma) * ((exp(lambda_L*iYear))^beta 
                                                  * iLabor^beta
                                                  * iCapStk^(1-beta) )^((sigma-1.0)/sigma) 
                          + gamma_E^(1.0/sigma)*(exp(lambda_E*iYear) 
                                                  * iEToFit)^((sigma-1.0)/sigma))^(sigma/(sigma-1.0))
  modelCES <- nlxb(model,
                   data=dataTable,
                   start=list(sigma=sigmaGuess, gamma_E=gamma_EGuess, beta=betaGuess, 
                              lambda_L=lambda_LGuess, lambda_E=lambda_EGuess),
                   trace=TRUE
                  )
  return(modelCES)
}
print("US Q")
print(summary(cesModel("US", "Q")))

@@

\citet{Stern:2012ey} developed a CES model of the following form

@@

\bibliographystyle{model2-names}
%%\bibliography{<your-bib-database>}
\bibliography{EconGrowth.bib}
